---
title: "Perth House Prices"
subtitle: "Practicing Multiple Linear Regression for the Coursera Machine Learning Specialisation"
output:
  html_document:
    theme:
      bg: "#ECF0F4"
      fg: "#202C39"
      primary: "#FF4F5C"
      base_font:
        google: "Sofia Sans"
      heading_font:
        google: "Sofia Sans"
      code_font:
        google: "Roboto Mono"
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
---

```{r setup, include=FALSE}
# Knitr chunk options
knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
```

Notebook prepared by [Clare Gibson](https://www.datatranslator.co.uk).

# Introduction

## Goals
In this notebook I will:

- practise optimizing $\vec{w}$ and $b$ using gradient descent
- practise what I have learned in weeks 1 and 2 of the [Stanford Machine Learning specialization](https://www.coursera.org/specializations/machine-learning-introduction?utm_medium=sem&utm_source=gg&utm_campaign=b2c_emea_x_multi_ftcof_career-academy_cx_dr_bau_gg_pmax_gc_s1_en_m_hyb_23-12_x&campaignid=20858198824&adgroupid=&device=c&keyword=&matchtype=&network=x&devicemodel=&creativeid=&assetgroupid=6490027433&targetid=&extensionid=&placement=&gad_source=1&gbraid=0AAAAADdKX6YAM0bxZ5FcdHU9SI3k0eeUu&gclid=CjwKCAiAiOa9BhBqEiwABCdG8xkDREktjtyaZgZG8005O_KWaT41YjYn2RSJAvqmzF1d84YTLnaA1BoCMpsQAvD_BwE)
- practise using R to implement a multiple linear regression model

## Tools
In this notebook I use the following tools:

- `here` for file location
- `tidyverse` for data wrangling
- `ggplot2` for plotting

```{r load-packages}
# Load packages
library(here)
library(tidyverse)
library(ggplot2)
```

## Data
In this notebook I use the [Kaggle Perth House Prices](https://www.kaggle.com/datasets/syuzai/perth-house-prices) dataset. The chunk below calls a script that loads the data into R and prepares it for the linear regression exercise.

```{r load-data}
# Load data
source(here("R/utils.R"))

# Store data in variable df
df <- model_1

# Show the head
head(df)
```

# Problem statement
In this exercise I am working on predicting house prices using various features. The training dataset contains `r nrow(df)` examples with three features (size, bedrooms and age). I will build a linear regression model using these values so that I can predict the price for other houses.

The code chunk below creates the `x_train` and `y_train` variables.

```{r create-variables}
# Create x_train and y_train
x_train <- as.matrix(df[,-1])
y_train <- as.matrix(df[,1])
```

The example data is now stored in matrix `x_train`. Each row of the matrix represents one training example. When you have $m$ training examples and $n$ features, `x_train` is a matrix with dimensions $(m,n)$. In this example the matrix has dimensions `r dim(x_train)`.

```{r view-dimensions}
dim(x_train)
```

