---
title: "Perth House Prices"
subtitle: "Practicing Multiple Linear Regression for the Coursera Machine Learning Specialisation"
output:
  html_document:
    theme:
      bg: "#ECF0F4"
      fg: "#202C39"
      primary: "#FF4F5C"
      base_font:
        google: "Sofia Sans"
      heading_font:
        google: "Sofia Sans"
      code_font:
        google: "Roboto Mono"
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
---

```{r setup, include=FALSE}
# Knitr chunk options
knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
```

Notebook prepared by [Clare Gibson](https://www.datatranslator.co.uk).

# Introduction

## Goals
In this notebook I will:

- practise optimizing $\vec{w}$ and $b$ using gradient descent
- practise what I have learned in weeks 1 and 2 of the [Stanford Machine Learning specialization](https://www.coursera.org/specializations/machine-learning-introduction?utm_medium=sem&utm_source=gg&utm_campaign=b2c_emea_x_multi_ftcof_career-academy_cx_dr_bau_gg_pmax_gc_s1_en_m_hyb_23-12_x&campaignid=20858198824&adgroupid=&device=c&keyword=&matchtype=&network=x&devicemodel=&creativeid=&assetgroupid=6490027433&targetid=&extensionid=&placement=&gad_source=1&gbraid=0AAAAADdKX6YAM0bxZ5FcdHU9SI3k0eeUu&gclid=CjwKCAiAiOa9BhBqEiwABCdG8xkDREktjtyaZgZG8005O_KWaT41YjYn2RSJAvqmzF1d84YTLnaA1BoCMpsQAvD_BwE)
- practise using R to implement a multiple linear regression model

## Tools
In this notebook I use the following tools:

- `here` for file location
- `tidyverse` for data wrangling
- `ggplot2` for plotting

```{r load-packages}
# Load packages
library(here)
library(tidyverse)
library(ggplot2)
```

## Data
In this notebook I use the [Kaggle Perth House Prices](https://www.kaggle.com/datasets/syuzai/perth-house-prices) dataset. The chunk below calls a script that loads the data into R and prepares it for the linear regression exercise.

```{r load-data}
# Load data
source(here("R/utils.R"))

# Store data in variable df
df <- model_1

# Show the head
head(df)
```

```{r load-alternative-data, eval=FALSE, include=FALSE}
# This data is the example used in the lab
df <- tibble(
  price = c(460, 232, 178),
  size = c(2104, 1416, 852),
  bedrooms = c(5, 3, 2),
  floors = c(1, 2, 1),
  age = c(45, 40, 35)
)

head(df)
```


# Problem statement
In this exercise I am working on predicting house prices using various features. The training dataset contains `r nrow(df)` examples with three features (size, bedrooms and age). I will build a linear regression model using these values so that I can predict the price for other houses.

## Setting the variables
The code chunk below creates the `x_train` and `y_train` variables.

```{r create-variables}
# Create x_train and y_train
x_train <- as.matrix(df[,-1])
y_train <- as.matrix(df[,1])
```

The example data is now stored in matrix `x_train`. Each row of the matrix represents one training example. When you have $m$ training examples and $n$ features, `x_train` is a matrix with dimensions $(m,n)$. In this example the `x_train` matrix has dimensions `r dim(x_train)`. The `y_train` matrix has dimensions `r dim(y_train)`.

```{r view-dimensions}
dim(x_train)
dim(y_train)
```

$\mathbf{w}$ is a vector with $n$ elements.

- Each element contains the parameter associated with one feature.
- in our dataset, n is 4.
- notionally, we draw this as a column vector

$$
\mathbf{w} = \begin{pmatrix}
w_0 \\ 
w_1 \\
\cdots\\
w_{n-1}
\end{pmatrix}
$$

$b$ is a scalar parameter.

For demonstration, $w$ and $b$ will be loaded with some initial selected values that are near the optimal.

```{r initialize-w-and-b}
# Initialise values for w and b
b_init <- 785.1811367994083
w_init <- c(0.39133535, 18.75376741, -26.42131618)
#w_init <-  c(0.39133535, 18.75376741, -53.36032453, -26.42131618)

b_init
w_init
```

# Predictions
## Using a for loop
One method to obtain a prediction is to loop over each feature, multiplying $w$ and $x$ together and then adding $b$ at the end.

```{r func-to-predict-with-for-loop}
# Predicting with for loop
predict_single_loop <- function(x, w, b) {
  n <- dim(x_train)[2]
  p <- 0
  
  for (i in 1:n) {
    p_i <- x[i] * w[i]
    p <- p + p_i
  }
  p <- p + b
  return(p)
}
```

I can use this function to obtain a prediction.
```{r predict-with-for-loop}
# Get a row from the training data
x_vec <- x_train[1,]
x_vec

# Make a prediction
f_wb <- predict_single_loop(x_vec, w_init, b_init)
paste0("Prediction: ", f_wb)
```

## Using vectorisation
An alternative mthod to obtain a prediction is to use vector operations by finding the dot product of $\vec{w}$ and $\vec{x}$.

```{r func-to-predict-with-vectorisation}
# Predicting with vectorisation
predict <- function(x, w, b) {
  p <- sum(x*w) + b
  return(p)
}
```

I can use this function to obtain a prediction.
```{r predict-with-vectorisation}
# Make a prediction
f_wb <- predict(x_vec, w_init, b_init)
paste0("Prediction: ", f_wb)
```

# Compute cost
The equation for the cost function with multiple variables $J(\mathbf{w},b)$ is:

$$
J(\mathbf{w},b) = \frac{1}{2m} \sum\limits_{i = 0}^{m-1} (f_{\mathbf{w},b}(\mathbf{x}^{(i)}) - y^{(i)})^2
$$ 

where:
$$
f_{\mathbf{w},b}(\mathbf{x}^{(i)}) = \mathbf{w} \cdot \mathbf{x}^{(i)} + b
$$ 
```{r func-to-compute-cost}
# Function to compute cost
compute_cost <- function(x, y, w, b) {
  m <- dim(x_train)[1]
  cost <- 0.0
  
  for (i in 1:m) {
    f_wb_i <- sum(x[i,] * w) + b
    cost <- cost + (f_wb_i - y[i]) ^ 2
  }
  cost <- cost / (2 * m)
  return(cost)
}
```

```{r compute-cost}
# Compute and display cost using our pre-chosen optimal parameters.
cost <- compute_cost(x_train, y_train, w_init, b_init)
paste0("Cost at optimal w: ", cost)
```

# Gradient descent
## Compute gradient
I will define a function to compute the gradients needed for gradient descent.

```{r func-to-compute-gradient}
# Compute the gradients
compute_gradient <- function(x, y, w, b) {
  m <- dim(x)[1]
  n <- dim(x)[2]
  dj_dw <- matrix(0,,n)
  dj_db <- 0.0
  
  for (i in 1:m) {
    err <- (sum(x[i,] * w) + b) - y[i,]
    for (j in 1:n) {
      dj_dw[j] <- dj_dw[j] + err * x[i, j]
    }
    dj_db <- dj_db + err
  }
  dj_dw <- dj_dw / m
  dj_db <- dj_db / m
  
  return(list(dj_db = dj_db, dj_dw = dj_dw))
}

# Compute and display gradient
tmp_grads <- compute_gradient(x_train, y_train, w_init, b_init)
tmp_grads
```

## Run gradient descent
I will write a function to perform the gradient descent algorithm to convergence.

```{r func-to-perform-gradient-descent}
# Gradient descent
gradient_descent <- function(x, y, w_in, b_in, gradient_function, alpha, num_iters) {
  w <- w_in
  b <- b_in
  
  for (i in 1:num_iters) {
    grads <- gradient_function(x, y, w, b)
    
    w <- w - alpha * grads$dj_dw
    b <- b - alpha * grads$dj_db
  }
  return(list(w = w, b = b))
}
```

In the next code chunk I will run the gradient descent function.

```{r run-gradient-descent}
# Initialise parameters
#initial_w <- c(0.0, 0.0, 0.0, 0.0)
initial_w <- c(0.0, 0.0, 0.0)
initial_b <- 0.0

# Gradient descent settings
iterations <- 1000
alpha <- 5.0e-7

# Run gradient descent
wb <- gradient_descent(
  x_train,
  y_train,
  initial_w,
  initial_b,
  compute_gradient,
  alpha,
  iterations
)

paste0("b found by gradient descent: ", round(wb$b,2))
paste0("w found by gradient descent: ", round(wb$w,2))

w <- wb$w
b <- wb$b
m = min(5, dim(x_train)[1])
for (i in 1:m) {
  print(paste0("Prediction: ", round(sum(x_train[i,] * w) + b, 2), " Target: ", y_train[i]))
}
```

